---
title: "IP WEEK 13"
author: "EZRA MAGETO"
date: "7/10/2021"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# DEFINING THE QUESTION

### Kira Plastinina is a Russian brand that is sold through a defunct chain of retail stores in Russia, Ukraine, Kazakhstan, Belarus, China, Philippines, and Armenia. The brand's Sales and Marketing team would like to understand their customer's behavior from data that they have collected over the past year. More specifically, they would like to learn the characteristics of customer groups.

------------------------------------------------------------------------

# METRICS OF SUCCESS

### 1.Perform clustering stating insights drawn from your analysis and visualizations.

### 2.Upon implementation, provide comparisons between the approaches learned this week i.e. K-Means clustering vs Hierarchical clustering highlighting the strengths and limitations of each approach in the context of your analysis.

# EXPERIMENTAL DESIGN 

### You will create a Markdown which will comprise the following sections.

------------------------------------------------------------------------

#### 1.Problem Definition

#### 2.Data Sourcing

#### 3.Check the Data

#### 4.Perform Data Cleaning

#### 5.Perform Exploratory Data Analysis (Univariate, Bivariate & Multivariate)

#### 6.Implement the Solution

#### 7.Challenge the Solution

#### 8.Follow up Questions

# 1.0 READING THE DATASET

```{r}
ecom<-read.csv("http://bit.ly/EcommerceCustomersDataset")
```

```{r}
##Previewing the top of the dataset
head(ecom)
```

```{r}
##Previewing the bottom of the data

tail(ecom)
```

# 1.1 EXPLORING THE DATASET

```{r}
#EXPLORING THE DATASET
##Viewing the dataset
View(ecom)
```

```{r}
##Checking the shape of the dataset
dim(ecom)
```

the dataset has 12330 rows and 18 columns

```{r}
##Checking information of the dataset

str(ecom)
```

```{r}
##checking the column names of the dataset

colnames(ecom)
```

```{r}
##Checking the datatypes of the dataset
class(ecom)
sapply(ecom,class)
```

```{r}
##Important features
ecomm <- ecom[, c(1,2,3,4,5,6,7,8,9,10)]
names(ecomm)
```

# 2.0 CLEANING THE DATASET

```{r}
#DATA CLEANING 
##Checking for missing values

sum(is.na(ecomm))
```

```{r}
##Dealing with missing values
ecomm<-na.omit(ecomm)
sum(is.na(ecomm))
```

```{r}
##Checking for duplicates
sum(duplicated(ecomm))
```

```{r}
##Dealing with duplicates
###To deal with duplicates i pick the non duplicated data
ecomm = ecomm[!duplicated(ecomm), ]
dim(ecomm)


```

```{r}
##Checking for outliers
product_related<-ecomm$ProductRelated
product_related_duration<-ecomm$ProductRelated_Duration
page_values<-ecomm$PageValues
Administrative<-ecomm$Administrative
Administrative_Duration<-ecomm$Administrative_Duration
Informational<-ecomm$Informational
Informational_Duration<-ecomm$Informational_Duration
BounceRates<-ecomm$BounceRates
ExitRates<-ecomm$ExitRates
SpecialDay<-ecomm$SpecialDay
boxplot(product_related,product_related_duration,page_values,SpecialDay,ExitRates,BounceRates,Informational_Duration,Informational,Administrative_Duration,Administrative)
```

the dataset contains some outliers in the Product Related Duration
feature

# 2.1 EXPLORATORY DATA ANALYSI

```{r}
#EXPLORATORY DATA ANALYSIS
##Univariate Analysis
####Summary of the descriptive statistics
summary(ecom)
```

From the descriptive statistics weekdays had the most numbers visiting
the site Many people who visited the site did not generate any revenue
for the brand. Only 1908 out of 12330 contributed to revenue addition
for the brand.

```{r}
#Plotting Histogram of selected features
library(DataExplorer)
plot_histogram(ecomm)
```

there is a positive skew in the distribution Of the features

```{r}
#creating bar plots for all categorical column
month_table <- table(ecom$Month)
os_table <- table(ecom$OperatingSystems)
browser_table <- table(ecom$Browser)
region_table <- table(ecom$Region)
traffic_table <- table(ecom$TrafficType)
visitor_table <- table(ecom$VisitorType)
weekend_table <- table(ecom$Weekend)
revenue_table <- table(ecom$Revenue)
```

```{r}
#creating barplot for month
barplot(month_table)
##the month of May had the most visits to the site while February was the least visited
```

```{r}
#Barplot indicating visitor type
barplot(visitor_table)
```

returning visitors form the majority of of the visitor type

```{r}
#weekend barplot
barplot(weekend_table)
```

the site is mostly visited during the weekdays

```{r}
#revenue barplot
barplot(revenue_table)
```

individuals who visit the site and generate revenue are less.Majority of
the people visit the site and don't generate any revenue

```{r}
#barplot on traffictype
barplot(traffic_table)
```

type 2 traffic is the highest followed by type 1 traffic

```{r}
#bar on region
barplot(region_table)
```

Region 1 has the majority of people who visit the site while region 9 is
the lowest

```{r}
#Bivariate Analysis 
##getting the covariance
page_values<-ecomm$PageValues
product_related<-ecomm$ProductRelated

cov(page_values,product_related)
```

```{r}
##correlation
page_values<-ecomm$PageValues
product_related<-ecomm$ProductRelated
cor(page_values,product_related)
```

```{r}
#correlation plot
library(ggcorrplot)
library(dplyr)
corr = round(cor(select_if(ecomm, is.numeric)), 4)
ggcorrplot(corr, hc.order = T, ggtheme = ggplot2::theme_grey,
   colors = c("cyan", "peachpuff4", "pink"), lab = T)

```

there is a high correlation between the bounce rate and the exit rate

```{r}
##plotting revenue per month

rev_month <- table(ecom$Revenue, ecom$Month)
barplot(rev_month, main = "Revenue per Month", col = c("black", "cyan"), beside = TRUE, 
        legend = rownames(rev_month), xlab = "Month")
```

November returned the highest number of revenue

```{r}
##plotting revenue per region

rev_region <- table(ecom$Revenue, ecom$Region)
barplot(rev_region, main = "Revenue per Region", col = c("black", "cyan"), beside = TRUE, 
        legend = rownames(rev_region), xlab = "Region")
```

region 1 returned more revenue with region 5 being the least

# 3.0 IMPLEMENTING THE SOLUTION

## K-MEANS CLUSTERING

#### Normalizing the required features

```{r}
normalize <- function(x){
  return ((x-min(x)) / (max(x)-min(x)))
}
ecomm$ProductRelated<- normalize(ecomm$ProductRelated)
ecomm$ProductRelated_Duration<- normalize(ecomm$ProductRelated_Duration)
ecomm$PageValues<- normalize(ecomm$PageValues)

head(ecomm)
```

```{r}
####selecting necesarry columns
ecomm1<-ecomm[c(5,6,9)]
colnames(ecomm1)
```

```{r}
#

# Applying the K-means clustering algorithm with number of centroids(k)=3
result<- kmeans(ecomm1,3) 
# Previewing the no. of records in each cluster
result$size 
```

```{r}
#value of cluster center datapoint
result$centers
```

```{r}
#cluster vector showing each record
result$cluster
```

```{r}
#clustering result visualization
par(mfrow = c(1,2), mar = c(5,4,2,2))
#plotting to see how product related duration and product related have been distributed
plot(ecomm1[,1:2], col = result$cluster) 
```

```{r}
#verifying the results of clustering
par(mfrow = c(2,2), mar = c(5,4,2,2))
#plotting product related duration and product related
plot(ecomm1[c(1,2)], col = result$cluster)
```

```{r}
#plotting page values and product related duration to check the distribution
plot(ecomm1[c(2,3)], col = result$cluster)
```

```{r}
#implementing with class attribute 'revenue'
ecomm1.class<- ecom[, "Revenue"]
```

```{r}
plot(ecomm1[c(2,3)], col = ecomm1.class)
```

```{r}
#HIERARCHICAL CLUSTERING
#scaling the data
ecomm1<-scale(ecomm1)
head(ecomm1)
```

```{r}
#computing the euclidean distance
ec<-dist(ecomm1,method = "euclidean")
#performing hierarchical clustering
HC<-hclust(ec,method = "ward.D2")
#plotting the dendrogram
plot(HC,cex=0.6,hang= -1)
```

# 4.0 CONCLUSION

### Majority of the people visit the site during the weekdays

### Majority of people visit the site but don't generate revenue

### Most of the returning visitors generate revenue to the brand,this is a good indication to have loyal customers

# 5.0 RECOMMENDATION

### Provide incentives to returning visitors to attract more new visitors
