---
output:
  html_document: default
  word_document: default
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

+-----------------------------------------------------------------------+
| title: "IP WEEK 12" author: "EZRA MAGETO" date: "7/5/2021" output:    |
| html_document                                                         |
+-----------------------------------------------------------------------+

# Defining the question

## Identifying which individuals are most likely to click on the cryptography course advert

# Defining the metrics for success

## Bi variate and Uni variate Exploratory data Analysis

# Understanding the context

## Target market for the cryptography course

# Experimental Design

```{r}
#Importing the data
df<-read.csv("http://bit.ly/IPAdvertisingData")
head(df)
```

```{r}
#previewing the dataset
View(df)
```

```{r}
#checking the column names
colnames(df)
```

```{r}
#previewing the head of the dataset
head(df)
```

```{r}
#previewing the tail of the dataset
tail(df)
```

```{r}
#structure of the data
str(df)
```

```{r}
#shape of the dataset
dim(df)
#the dataset has 1000 rows and 10 columns
```

# Data cleaning

## Checking for Missing values

```{r}
sum(is.na(df))
#there are no missing values in the dataset
```

# Duplicates

```{r}
#checking for duplicates
sum(duplicated(df))
#there are no duplicates in the dataset
```

```{r}
#removal of incomplete objects in the dataset
head(na.omit(df))
```

# Checking for outliers in area . income column

```{r}
#boxplots

boxplot(df$Area.Income)
```

# Uni-variate Analysis

## Measures of central Tendency

```{r}
summary(df)

```

# Finding the mode of age

```{r}
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
df.Age.mode <- getmode(df$Age)
df.Age.mode
```

```{r}
#finding the median of age
df.Age.median <- median(df$Age)
df.Age.median 
#[1] 35 is the median age
```

# Measures of dispersion

### selecting the most important columns for calculating measures of dispersion-AGE and DAILY INTERNET USAGE

### getting the minimum for internet usage

```{r}
df.Daily.Internet.Usage.min <- min(df$Daily.Internet.Usage)
df.Daily.Internet.Usage.min
#[1] 104.78
```

```{r}
#finding the minimum for age
df.Age.min <- min(df$Age)
df.Age.min
#[1] 19 is the min age
```

```{r}
#Maximum
df.Daily.Internet.Usage.max <- max(df$Daily.Internet.Usage)
df.Daily.Internet.Usage.max
#[1] 269.96 is the maximum internet usage
```

```{r}
#maximum age
df.Age.max <- max(df$Age)
df.Age.max
#[1] 61 is the maximum age
```

```{r}
#range of internet usage
df.Daily.Internet.Usage.range <- range(df$Daily.Internet.Usage)
df.Daily.Internet.Usage.range
#[1] 104.78 269.96 is the range of internet usage
```

```{r}
#range of age
df.Age.range <- range(df$Age)
df.Age.range
#[1] 19 61 is the range of age
```

```{r}
# Finding the quantile of internet usage
df.Daily.Internet.Usage.quantile <- quantile(df$Daily.Internet.Usage)
df.Daily.Internet.Usage.quantile
   #   0%      25%      50%      75%     100% 
#104.7800 138.8300 183.1300 218.7925 269.9600 
```

```{r}
#Quantile of age
df.Age.quantile <- quantile(df$Age)
df.Age.quantile
# 0%  25%  50%  75% 100% 
#  19   29   35   42   61
```

```{r}
#variance of internet usage
df.Daily.Internet.Usage.variance <- var(df$Daily.Internet.Usage)
df.Daily.Internet.Usage.variance
#[1] 1927.415 is the variance of internet usage
```

```{r}
#variance of age
df.Age.variance <- var(df$Age)
df.Age.variance
#[1] 77.18611 is the variance of age
```

```{r}
#getting the standard deviation
df.Daily.Internet.Usage.sd <- sd(df$Daily.Internet.Usage)
df.Daily.Internet.Usage.sd
#[1] 43.90234 is the standard deviation of internet usage
```

```{r}
#standard deviation of age
df.Age.sd <- sd(df$Age)
df.Age.sd

```

# Uni-variate plots

```{r}
#getting outliers in the column Daily Time Spent on Site
boxplot(df$Daily.Time.Spent.on.Site)
```

```{r}
# getting  columns
Daily.Internet.Usage <- df$Daily.Internet.Usage

#  the frequency distribution
Daily.Internet_Usage.frequency <- table(Daily.Internet.Usage)
```

```{r}

# plotting the bargraph
barplot(Daily.Internet.Usage)
```

```{r}
#ploting bar plot on male column
gender <- df$Male
gender_freq <- table(gender)
barplot(gender_freq,
        col = "red")
 #0 represented slightly more people compared to 1       
```

# histogram of daily internet usage

```{r}
hist(df$Daily.Internet.Usage)
#there is a high frequency between 100 and 150 in daily internet usage
```

# histogram of age

```{r}
hist(df$Age)

#majority of the people fall between 25-40 years
```

## Bi variate Analysis

```{r}
timespent <- df$Daily.Time.Spent.on.Site
internetusage<- df$Daily.Internet.Usage

# Using the cov() function to determine the covariance

cov(timespent, internetusage)
```

```{r}
# finding the correlation

Age<-df$Age
Daily.Internet.Usage<-df$Daily.Internet.Usage
#using the cor function to find the correlation
cor(Age, Daily.Internet.Usage)
#There is a negative correlation between age and daily internet usage
```

# creating a scatter-plot

```{r}
timespent <- df$Daily.Time.Spent.on.Site
internetusage<- df$Daily.Internet.Usage

plot(timespent, internetusage, xlab="Daily time spent on site", ylab="Daily Internet Usage")
```

# SUPERVISED LEARNING

\**\#\#Decision tree model*\#

```{r}
library(partykit)

```

```{r}
# Creating the input data frame.
input.dat <- df

# Creating the tree.
  output.tree <- ctree( Daily.Internet.Usage ~ Age + Clicked.on.Ad + 
Daily.Time.Spent.on.Site, 
  data = input.dat)
```

\#\#Plotting the tree

```{r}
plot(output.tree)
```



# CONCLUSION

### It was realized that the higher the age, the lower the time that was spent on the site. There were no outliers in the data which made it easier for analysis The range in age of people who visited the site was between 19 an 61 years

# Recommendation

### More focus should be emphasized on the young people compared to the older people
